---
title: "Lab2 Saunders"
author: "Taylor Saunders"
date: "2023-11-05"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
library(ggplot2)
library(ggplot2movies)
data(movies)
```


```{r}
library(dplyr)
```

Question 1

```{r}
min(movies$year)
max(movies$year)
print("The range of years of production of the movies of this dataset is 1893 to 2005" )
```

Question 2

```{r}
#part 1
has_budget <- sum(is.na(movies$budget))
print(has_budget) 
#dim(movies) #checking values are reasonable
no_budget <- sum(!is.na(movies$budget))
print(no_budget)
has_budget_percent <- 53573 / 58788  * 100
print(has_budget_percent)

no_budget_percent <- 5215 / 58788 * 100
print(no_budget_percent)
```

91% of the movies had a value for their budget, while 9% of the movies did not have a listed budget

```{r}
#part 2
top_5_expense <- arrange(movies, desc(budget)) #arrange in descending order based on length value
head(top_5_expense, n = 5)
```
The top 5 most expensive movies in this data set are Spider-Man 2 and the Titanic, both with the same budget of 200000000, Troy, Terminator 3: Rise of the Machines, and Waterworld.

Question 3

```{r}
top_5 <- arrange(movies, desc(length)) #arrange in descending order based on length value to see highest value
head(top_5, n = 5)
```
The top 5 longest movies are The Cure for Insomnia, The Longest Most Meaningless Movie in the World, Four Stars, Resan, and Out 1. 


Question 4

```{r}
shortest_movie <- arrange(movies, length) #sort in ascending order
head(shortest_movie)
table(shortest_movie$length == 1) #more than one movie has the shortest length
```
The head of the first 5 shortest movies is 17 seconds to Sophie, 2 A.M. in the Subway, Admiral Cigarette, Admiral Dewey Leading Land Parade, Alphonse and Gaston No. 3.
However, there are 169 movies in this data set with the length "1".

Question 5

```{r}
genre <- movies %>% #dataset being used
  select(Action, Animation, Comedy, Drama, Documentary, Romance, Short) %>% #the genres for our plot
  colSums() #sum of each column

barplot(genre,
        main = "Number of Movies by Genre",
        xlab = "Genre",
        ylab = "Count of Movies",
        cex.names=0.7,
        col = "red"
)
```
Question 6
```{r}
 
 action = filter(movies, Action == 1)
 actionrating = mean(action$rating)
 
 animation = filter(movies, Animation == 1)
 animationrating = mean(animation$rating)
 
 comedy = filter(movies, Comedy == 1)
 comedyrating = mean(action$rating)
 
 drama = filter(movies, Drama == 1)
 dramarating = mean(drama$rating)
 
 Doc = filter(movies, Documentary == 1)
 docrating = mean(Doc$rating)
 
 shorts = filter(movies, Short == 1)
 shortrating = mean(shorts$rating)
 
 romance = filter(movies, Romance == 1)
 romancerating = mean(romance$rating)
 
 averagerate <- c(actionrating, animationrating, comedyrating, docrating, 
               dramarating, shortrating, romancerating) 
 
 counts = as.vector(averagerate)
 xx <-
   barplot(
     averagerate,
     main = "Average rate of all movies",
     names = c('Action','Animation','Comedy','Documentary','Drama','Short','Romance'),
     xlab = "Genre",
     ylim = c(0, max(averagerate) + 4),
     cex.names=0.7,
     col = "Red"
   )
 text(
   x <- xx,
   y = averagerate + 1,
   label = as.character(averagerate),
   cex = 0.5,
   col = "Black"
 )
```
Question 7
```{r}
 action = filter(movies, Action == 1, year >= 2000 & year <= 2005)
 actionrating = mean(action$rating)
 
 animation = filter(movies, Animation == 1, year >= 2000 & year <= 2005)
 animationrating = mean(animation$rating)
 
 comedy = filter(movies, Comedy == 1, year >= 2000 & year <= 2005)
 comedyrating = mean(action$rating)
 
 drama = filter(movies, Drama == 1, year >= 2000 & year <= 2005)
 dramarating = mean(drama$rating)
 
 Doc = filter(movies, Documentary == 1, year >= 2000 & year <= 2005)
 docrating = mean(Doc$rating)
 
 shorts = filter(movies, Short == 1, year >= 2000 & year <= 2005)
 shortrating = mean(shorts$rating)
 
 romance = filter(movies, Romance == 1, year >= 2000 & year <= 2005)
 romancerating = mean(romance$rating)
 
 averagerate <- c(actionrating, animationrating, comedyrating, docrating, 
               dramarating, shortrating, romancerating) 
 
 counts = as.vector(averagerate)
 xx <-
   barplot(
     averagerate,
     main = "Average rate of all movies",
     names = c('Action','Animation','Comedy','Documentary','Drama','Short','Romance'),
     xlab = "Genre",
     ylim = c(0, max(averagerate) + 4),
     cex.names=0.6,
     col = "Red"
   )
 text(
   x <- xx,
   y = averagerate + 1,
   label = as.character(averagerate),
   cex = 0.5,
   col = "Black"
 )
```

Question 8

```{r}
selected_genres <- movies %>%
  select(Action, Animation, Comedy, Drama, Documentary, Romance, year) %>%
  filter(year >= 1990)

plot_top6 <- function(genre){
  plot_genres <- selected_genres %>%
  group_by(year) %>%
  summarise(
    action = sum(Action),
    animation = sum(Animation),
    comedy = sum(Comedy),
    drama = sum(Drama),
    documentary= sum(Documentary),
    romance = sum(Romance),
  )
 
}
print(plot_top6())

```
Comment: When using a function, the appropriate tables are made. However, I am uncertain as to how to graph them using the function. 

Plotting without the use of a function: 
```{r}
plot_genre <- 
  ggplot2movies::movies %>%
  filter(year >= 1990) %>%
  select(Action, Animation, Comedy, Drama, Documentary, Romance, year) %>%
  group_by(year) %>%
  summarise(
    action = sum(Action),
    animation = sum(Animation),
    comedy = sum(Comedy),
    drama = sum(Drama),
    documentary = sum(Documentary),
    romance = sum(Romance),
  )
 ggplot(plot_genre, aes(x = year)) +
  geom_line(aes(y = action, color = "action")) + 
  geom_line(aes(y = animation, color = "animation")) + 
  geom_line(aes(y = comedy, colur = "comedy")) + 
  geom_line(aes(y = drama, color = "drama")) +
  geom_line(aes(y = documentary, color = "doc")) +
  geom_line(aes(y = romance, color = "romance")) +
  ylab("count")
```


Question 9
1. How many movies were published in 2002?
2. What is the top movie for the Action Genre in 2002?
3. What is the top movie for the Action Genre in 2002?

```{r}
#1
#3
movies_2002 <- sum(movies$year == "2002")
print(movies_2002)
```
There were 2,168 movies published in 2002

```{r}
#2
top_action <- filter(movies, Action == 1, year == 2002)
top_action <- arrange(top_action, desc(rating)) 
head(top_action)
```
The top rated action movie in 2002 was Sundown, with a rating of 9.5.

```{r}
#3
top_drama <- filter(movies, Drama == 1, year == 2002)
head(top_drama)
top_drama <- arrange(top_drama, desc(rating)) 
head(top_drama)
dim(top_drama)
```

The top rated drama movie in 2002 was Mutual Admiration Society, with a rating of 9.6. 


